---
title: 'Reproducible Research: Peer Assessment 1'
author: "AhmedShawky"
date: "May 1, 2018"
output: html_document
---

#Code for reading in the dataset and/or processing the data

```{r}
activity<-read.csv("activity.csv")
summary(activity)
```

#Histogram of the total number of steps taken each day

```{r}
num_step_per_day <- with(activity, tapply(steps, as.factor(activity$date), sum, na.rm = T))
hist(num_step_per_day,xlab='Total Steps per day',main ='Histogram of the total number of steps taken each day')
summary(num_step_per_day)

mean(num_step_per_day, na.rm = TRUE)
```

#What is the average daily activity pattern?

##time series plot
```{r}
activity_Without_NA <- activity[!is.na(activity$steps),]
avg_steps <- with(activity_Without_NA, tapply(steps, activity_Without_NA$interval, mean))
plot(unique(activity$interval), avg_steps, type = "l", main = "time series plot", xlab = "interval", ylab = "avg steps")
```


##Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
```{r}
avg_steps[which.max(avg_steps)]
```


#Imputing missing values

##the total number of missing values

```{R}
length(activity[is.na(activity$steps),]$steps)
```

## Set the missing values with the mean value & Create New DataSet



```{R}
activity_fill_na <- activity;
activity_fill_na[is.na(activity_fill_na$steps),]$steps <-mean(avg_steps)

```

##how imputing missing value impact the dataset

```{r}
num_step_per_day_Fill_Missing_na <- with(activity_fill_na, tapply(steps, as.factor(activity$date), sum, na.rm = T))
hist(num_step_per_day_Fill_Missing_na,xlab='Total Steps per day',main ='Histogram of the total number of steps taken each day')

summary(num_step_per_day_Fill_Missing_na)
summary(num_step_per_day)
```

##Weekday

```{r}
#add new col "isweekend"
activity_fill_na$isweekend<-weekdays(as.Date(activity_fill_na$date)) %in% c('Sunday' ,'Saturday')
activity_fill_na[activity_fill_na$isweekend==T,]$DayType<-"Weekend"
activity_fill_na[activity_fill_na$isweekend==F,]$DayType<-"Weekday"


activity_weekend<-activity_fill_na[activity_fill_na$isweekend==T,]
activity_weekday<-activity_fill_na[activity_fill_na$isweekend==F,]

avg_steps_weekend <- with(activity_weekend, tapply(steps, activity_weekend$interval, mean))
avg_steps_weekday <- with(activity_weekday, tapply(steps, activity_weekday$interval, mean))

plot(unique(activity_weekday$interval), avg_steps_weekday, type = "l", main = "time series plot", xlab = "interval", ylab = "avg steps" ,col='red')
lines(unique(activity_weekend$interval),avg_steps_weekend,col="blue")

legend("topright",  legend=c('Weekend', 'Weekday'),col=c( 'blue', 'red'), lty=c( 1, 1))

```



